---
title: "DATA607 WK1 Homework: Mushrooms Dataset"
author: "Javern Wilson"
date: "January 28, 2019"
output: 
  html_document:
    code_folding: hide
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# {.tabset .tabset-fade .tabset-pills}

## Original Mushroom Dataset

At first glance it may be difficult to understand what the database is trying to convey.

```{r Original, echo=TRUE}

url = "https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data"

originalMdf <- read.csv(url, header = F, sep = ",", stringsAsFactors = F)
head(originalMdf, 10)

```

## List of Attributes 
For reference

1. Class
2. Cap Shape                          => **C_shape**
3. Cap Surface                        => **C_surface**
4. Cap Color                          => **C_color**
5. Bruises
6. Oder
7. Gill Attachment                    => **G_Attachment**
8. Gill Spacing                       => **G_spacing**
9. Gill Size                          => **G_Size**
10. Gill Color                        => **G_color**
11. Stalk Shape                       => **S_shape**
12. Stalk Root                        => **S_root**
13. Stalk Surface Above Ring          => **S_surfaceAR**
14. Stalk Surface Below Ring          => **S_surfaceBR**
15. Stalk Color Above Ring            => **S_colorAR**
16. Stalk Color Below Ring            => **S_colorBR**
17. Veil Type                         => **V_type**
18. Veil Color                        => **V_color**
19. Ring Number                       => **R_number**
20. Ring Type                         => **R_type**
21. Spore Print Color                 => **SporeCol**
22. Population
23. Habitat


## Modified Mushroom Dataset

A more detailed display of attributes that categorizes the different types of Mushrooms.

```{r modified, echo=TRUE}

Mdf <- originalMdf

#Renaming the column names for readability
colnames(Mdf) <- c("Class", "C_shape", "C_surface", "C_color", "Bruises", "Odor", "G_Attachment", "G_spacing", "G_size", "G_color", "S_shape", "S_root", "S_surfaceAR", "S_surfaceBR", "S_colorAR", "S_colorBR", "V_type", "V_color", "R_number", "R_type", "SporeCol", "Population", "Habitat")

#Renaming values in the column for readability
Mdf$Class[Mdf$Class == 'p'] <- "Poisonous"
Mdf$Class[Mdf$Class == 'e'] <- "Edible"

Mdf$C_shape[Mdf$C_shape == 'b'] <- "Bell"
Mdf$C_shape[Mdf$C_shape == 'c'] <- "Conical"
Mdf$C_shape[Mdf$C_shape == 'x'] <- "Convex"
Mdf$C_shape[Mdf$C_shape == 'f'] <- "Flat"
Mdf$C_shape[Mdf$C_shape == 'k'] <- "Knobbed"
Mdf$C_shape[Mdf$C_shape == 's'] <- "Sunken"

Mdf$C_surface[Mdf$C_surface == 'f'] <- "Fibrous"
Mdf$C_surface[Mdf$C_surface == 'g'] <- "Grooves"
Mdf$C_surface[Mdf$C_surface == 'y'] <- "Scaly"
Mdf$C_surface[Mdf$C_surface == 's'] <- "Smooth"

Mdf$C_color[Mdf$C_color == 'n'] <- "Brown"
Mdf$C_color[Mdf$C_color == 'b'] <- "Buff"
Mdf$C_color[Mdf$C_color == 'c'] <- "Cinnamon"
Mdf$C_color[Mdf$C_color == 'g'] <- "Gray"
Mdf$C_color[Mdf$C_color == 'r'] <- "Green"
Mdf$C_color[Mdf$C_color == 'p'] <- "Pink"
Mdf$C_color[Mdf$C_color == 'u'] <- "Purple"
Mdf$C_color[Mdf$C_color == 'e'] <- "Red"
Mdf$C_color[Mdf$C_color == 'w'] <- "White"
Mdf$C_color[Mdf$C_color == 'y'] <- "Yellow"

Mdf$Bruises[Mdf$Bruises == 't'] <- "Yes"
Mdf$Bruises[Mdf$Bruises == 'f'] <- "No"

Mdf$Odor[Mdf$Odor == 'a'] <- "Almond"
Mdf$Odor[Mdf$Odor == 'l'] <- "Anise"
Mdf$Odor[Mdf$Odor == 'c'] <- "Creosote"
Mdf$Odor[Mdf$Odor == 'y'] <- "Fishy"
Mdf$Odor[Mdf$Odor == 'f'] <- "Foul"
Mdf$Odor[Mdf$Odor == 'm'] <- "Musty"
Mdf$Odor[Mdf$Odor == 'n'] <- "None"
Mdf$Odor[Mdf$Odor == 'p'] <- "Pungent"
Mdf$Odor[Mdf$Odor == 's'] <- "Spicy"

Mdf$G_Attachment[Mdf$G_Attachment == 'a'] <- "Attached"
Mdf$G_Attachment[Mdf$G_Attachment == 'd'] <- "Descending"
Mdf$G_Attachment[Mdf$G_Attachment == 'f'] <- "Free"
Mdf$G_Attachment[Mdf$G_Attachment == 'n'] <- "Notched"

Mdf$G_spacing[Mdf$G_spacing == 'c'] <- "Closed"
Mdf$G_spacing[Mdf$G_spacing == 'w'] <- "Crowded"
Mdf$G_spacing[Mdf$G_spacing == 'd'] <- "Distant"

Mdf$G_size[Mdf$G_size == 'b'] <- "Broad"
Mdf$G_size[Mdf$G_size == 'n'] <- "Narrow"

Mdf$G_color[Mdf$G_color == 'k'] <- "Black"
Mdf$G_color[Mdf$G_color == 'n'] <- "Brown"
Mdf$G_color[Mdf$G_color == 'b'] <- "Buff"
Mdf$G_color[Mdf$G_color == 'h'] <- "Chocolate"
Mdf$G_color[Mdf$G_color == 'g'] <- "Gray"
Mdf$G_color[Mdf$G_color == 'r'] <- "Green"
Mdf$G_color[Mdf$G_color == 'o'] <- "Orange"
Mdf$G_color[Mdf$G_color == 'p'] <- "Pink"
Mdf$G_color[Mdf$G_color == 'u'] <- "Purple"
Mdf$G_color[Mdf$G_color == 'e'] <- "Red"
Mdf$G_color[Mdf$G_color == 'w'] <- "White"
Mdf$G_color[Mdf$G_color == 'y'] <- "Yellow"

Mdf$S_shape[Mdf$S_shape == 'e'] <- "Enlarging"
Mdf$S_shape[Mdf$S_shape == 't'] <- "Tapering"

Mdf$S_root[Mdf$S_root == 'b'] <- "Bulbous"
Mdf$S_root[Mdf$S_root == 'c'] <- "Club"
Mdf$S_root[Mdf$S_root == 'u'] <- "Cup"
Mdf$S_root[Mdf$S_root == 'e'] <- "Equal"

Mdf$S_surfaceAR[Mdf$S_surfaceAR == 'f'] <- "Fibrous"
Mdf$S_surfaceAR[Mdf$S_surfaceAR == 'y'] <- "Scaly"
Mdf$S_surfaceAR[Mdf$S_surfaceAR == 'k'] <- "Silky"
Mdf$S_surfaceAR[Mdf$S_surfaceAR == 's'] <- "Smooth"

Mdf$S_surfaceBR[Mdf$S_surfaceBR == 'f'] <- "Fibrous"
Mdf$S_surfaceBR[Mdf$S_surfaceBR == 'y'] <- "Scaly"
Mdf$S_surfaceBR[Mdf$S_surfaceBR == 'k'] <- "Silky"
Mdf$S_surfaceBR[Mdf$S_surfaceBR == 's'] <- "Smooth"
 
Mdf$S_colorAR[Mdf$S_colorAR == 'n'] <- "Brown"
Mdf$S_colorAR[Mdf$S_colorAR == 'b'] <- "Brown"
Mdf$S_colorAR[Mdf$S_colorAR == 'c'] <- "Brown"
Mdf$S_colorAR[Mdf$S_colorAR == 'g'] <- "Gray"
Mdf$S_colorAR[Mdf$S_colorAR == 'o'] <- "Orange"
Mdf$S_colorAR[Mdf$S_colorAR == 'p'] <- "Pink"
Mdf$S_colorAR[Mdf$S_colorAR == 'e'] <- "Red"
Mdf$S_colorAR[Mdf$S_colorAR == 'w'] <- "White"
Mdf$S_colorAR[Mdf$S_colorAR == 'y'] <- "Yellow"

Mdf$S_colorBR[Mdf$S_colorBR == 'n'] <- "Brown"
Mdf$S_colorBR[Mdf$S_colorBR == 'b'] <- "Brown"
Mdf$S_colorBR[Mdf$S_colorBR == 'c'] <- "Brown"
Mdf$S_colorBR[Mdf$S_colorBR == 'g'] <- "Gray"
Mdf$S_colorBR[Mdf$S_colorBR == 'o'] <- "Orange"
Mdf$S_colorBR[Mdf$S_colorBR == 'p'] <- "Pink"
Mdf$S_colorBR[Mdf$S_colorBR == 'e'] <- "Red"
Mdf$S_colorBR[Mdf$S_colorBR == 'w'] <- "White"
Mdf$S_colorBR[Mdf$S_colorBR == 'y'] <- "Yellow"

Mdf$V_type[Mdf$V_type == 'p'] <- "Partial"
Mdf$V_type[Mdf$V_type == 'u'] <- "Universal"

Mdf$V_color[Mdf$V_color == 'n'] <- "Brown"
Mdf$V_color[Mdf$V_color == 'o'] <- "Orange"
Mdf$V_color[Mdf$V_color == 'w'] <- "White"
Mdf$V_color[Mdf$V_color == 'y'] <- "Yellow"

Mdf$R_number[Mdf$R_number == 'n'] <- "None"
Mdf$R_number[Mdf$R_number == 'o'] <- "One"
Mdf$R_number[Mdf$R_number == 't'] <- "Two"

Mdf$R_type[Mdf$R_type == 'c'] <- "Cobwebby"
Mdf$R_type[Mdf$R_type == 'e'] <- "Evanescent"
Mdf$R_type[Mdf$R_type == 'f'] <- "Flaring"
Mdf$R_type[Mdf$R_type == 'l'] <- "Large"
Mdf$R_type[Mdf$R_type == 'n'] <- "None"
Mdf$R_type[Mdf$R_type == 'p'] <- "Pendant"
Mdf$R_type[Mdf$R_type == 's'] <- "Sheathing"
Mdf$R_type[Mdf$R_type == 'z'] <- "Zone"

Mdf$SporeCol[Mdf$SporeCol == 'k'] <- "Black" 
Mdf$SporeCol[Mdf$SporeCol == 'n'] <- "Brown"
Mdf$SporeCol[Mdf$SporeCol == 'b'] <- "Buff"
Mdf$SporeCol[Mdf$SporeCol == 'h'] <- "Chocolate"
Mdf$SporeCol[Mdf$SporeCol == 'r'] <- "Green"
Mdf$SporeCol[Mdf$SporeCol == 'o'] <- "Orange"
Mdf$SporeCol[Mdf$SporeCol == 'u'] <- "Purple"
Mdf$SporeCol[Mdf$SporeCol == 'w'] <- "White"
Mdf$SporeCol[Mdf$SporeCol == 'y'] <- "Yellow"

Mdf$Population[Mdf$Population == 'a'] <- "Abundant"
Mdf$Population[Mdf$Population == 'c'] <- "Clustered"
Mdf$Population[Mdf$Population == 'n'] <- "Numerous"
Mdf$Population[Mdf$Population == 's'] <- "Scattered"
Mdf$Population[Mdf$Population == 'v'] <- "Several"
Mdf$Population[Mdf$Population == 'y'] <- "Solitary"

Mdf$Habitat[Mdf$Habitat == 'g'] <- "Grasses"
Mdf$Habitat[Mdf$Habitat == 'l'] <- "Leaves"
Mdf$Habitat[Mdf$Habitat == 'm'] <- "Meadows"
Mdf$Habitat[Mdf$Habitat == 'p'] <- "Paths"
Mdf$Habitat[Mdf$Habitat == 'u'] <- "Urban"
Mdf$Habitat[Mdf$Habitat == 'w'] <- "Waste"
Mdf$Habitat[Mdf$Habitat == 'd'] <- "Woods"

head(Mdf)

```

## Subset

Displays a table of the characterics of Mushrooms that mycologists mostly observe 
```{r Exploration, echo=TRUE, message=FALSE, warning=FALSE}

library(plyr)
library(dplyr)
M2df <- select(Mdf, Class, C_shape, C_surface, C_color, Habitat)

head(M2df, 25)

#Totals for values in each column 
summary(factor(M2df$Class))
summary(factor(M2df$C_shape))
summary(factor(M2df$C_surface))
summary(factor(M2df$C_color))
summary(factor(M2df$Habitat))

```
<br/>

#### Characteristics of Poisonous mushrooms
```{r poisonous, echo=TRUE, message=FALSE, warning=FALSE}
#Details of poisonous mushrooms
poison <- M2df %>% filter(Class == 'Poisonous')
head(poison, 20)
table(poison$Habitat)

```
Most poisonous mushroom a found in the Woods.
<br/>


####Characteristics of Edible mushrooms
```{r edible, echo=TRUE}
#Details of edible mushrooms

edible <- M2df %>% filter(Class == 'Edible')
head(edible, 20)
table(edible$C_surface)
```
Fibrous Mushrooms are mostly edible.

### 