---
title: 'DATA607 WK7: HTML, XML, JSON'
author: "Javern Wilson"
date: "March 14, 2019"
output:
   html_document:
     theme: paper
     highlight: textmate
     toc: true
     code_folding: hide
     toc_float: true
     collapsed: false
     smooth_scroll: false
  
---

#### Task: Pick three of your favorite books on one of your favorite subjects.   At least one of the books should have more than one author.  For each book, include the title, authors, and two or three other attributes that you find interesting. Take the information that you've selected about these three books, and separately create three files which store the book's information in HTML (using an html table), XML, and JSON formats (e.g.  "books.html", "books.xml", and "books.json").   

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(XML)
library(jsonlite)
library(rjson)
library(stringr)
library(rvest)
library(RCurl)
library(dplyr)
library(kableExtra)
```

## HTML
```{r HTML}

html_file <- read_html("https://raw.githubusercontent.com/javernw/JWCUNYAssignments/master/books.html")

#parse html file into R
books_html <- htmlParse(html_file)
books_html

books_html_df <- books_html %>%  readHTMLTable(header = TRUE, stringAsFactors = FALSE) %>% as.data.frame()

#clean table
books_html_df <- books_html_df[,-6] #remove null column

#rename columns
names(books_html_df) <- c("Title", "Author(s)", "Edition", "Publisher", "Date Published")

kable(books_html_df) %>% kable_styling(bootstrap_options = "striped")

```
<br/> <br/>

## XML
```{r XML}
#get url
 xml_file <- getURL("https://raw.githubusercontent.com/javernw/JWCUNYAssignments/master/books.xml")
#extract information from web into R
books_xml <- xmlParse(xml_file)
books_xml

#estract root node
xroot <- xmlRoot(books_xml)

#convert books.xml to dataframe
books_xml_df <- xmlToDataFrame(xroot)
names(books_xml_df) <- c("Title", "Author(s)", "Edition", "Publisher", "Date Published")

kable(books_xml_df) %>% kable_styling(bootstrap_options = "striped")

```
<br/> <br/>

## JSON
```{r JSON}

json_file <- "https://raw.githubusercontent.com/javernw/JWCUNYAssignments/master/books.json"

#input file into R
books_json <- fromJSON(file = json_file) 
books_json 

#class(books_json) is a list

#convert "books_json" from a list to a dataframe
books_json_df <- data.frame(matrix(unlist(books_json), nrow=3, byrow=T),stringsAsFactors=FALSE)

names(books_json_df) <- c("Title", "Author(s)", "Edition", "Publisher", "Date Published")

kable(books_json_df) %>% kable_styling(bootstrap_options = "striped")

 

```
<br/> <br/> <br/>

## Conclusion
#### All dataframes turns out to be the same. 
