---
title: "DATA607 WK3 Homework: Regex"
author: "Javern Wilson"
date: "February 13, 2019"
output: 
  html_document:
        toc: true
        toc_float: true
        toc_depth: 2 
        code_folding: hide
        theme: journal
---

## Problem 3
## {.tabset .tabset-fade}

### Raw Data
```{r rawdata, message=FALSE, warning=FALSE}
library(stringr)
library(tidyverse)

raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555 -6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
raw.data

#Extracts all letters except phone numbers
rawnames <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}")) 
rawnames

```

### First and Last name
(a) Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name. 

```{r fname lname}
#Switch first name with last name with back-referencing
name <- sub("(\\w+),\\s((\\w\\.)|(\\w+))","\\2 \\1", rawnames)

#to fix this name up a bit
name <- str_replace(name, "C. Burns Montgomery", "C. Montgomery Burns")
name

#? An alternative would be greatly appreciated in outputting the first initial
first_name <- unlist(str_extract_all(name, "(([a-zA-Z]+ )|([A-Z]\\............ )|( [A-Za-z]+$))"))
first_name

#getting last names
last_name <- unlist(str_extract_all(name, "[^ ]+$"))
last_name

#concat first and last name
FLname <- str_c(first_name, last_name)
FLname

```

### Detect Title
(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.). 

```{r detect}
# Extract titles Rev. and Dr.
title <- unlist(str_extract_all(name, "\\w{2,}\\. "))

#checks to see if anyone has the title Rev. or Dr.
anytitle <- str_detect(name, title)
anytitle


```

### Detect Midname
(c) Construct a logical vector indicating whether a character has a second name. 

```{r}
#Extract character with second name
midname <- unlist(str_extract_all(name, "[A-Z]\\. \\w+\\s")) 

str_detect(name, midname ) #checks for persons with a second name

```

## Problem 4
## {.tabset .tabset-fade}

####Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression. 

### [0-9]+\\$ 
   - This expression matches a digit followed by one or more digits then a dollar ($) sign.
```{r}
test <- c("12335", "45566", "$", "4666$", "463211", "vvvvv4$", "sksgmskmg")

str_extract_all(test, "[0-9]+\\$")


```
   
   
### \\b[a-z]{1,4}\\b 
  - Matches up to a four letter string. Alpha lowercase characters only.
```{r}
library(stringi)

#randomly generates 5 strings with up to a length of 6
words <- tolower(stri_rand_strings(5, sample(1:6, 5, replace=TRUE)))
words
str_extract_all(words, "\\b[a-z]{1,4}\\b")

```


### .*?\\.txt$ 
 
  - This expression matches to a string with .txt at the end.
```{r}
sometext <- c("a.txt", "thisproject23.txt", "help.com")
sometext
str_extract_all(sometext, ".*?\\.txt$")
```


### \\d{2}/\\d{2}/\\d{4} 
  - The forward slash (/) was the giveaway for this one.This expression is matched to a date data. Digits only with the forward slash format.  

```{r}

dates <- c("02/22/2019", "ab/cd/efhi", "1a-55-4789", "02-22-2019")
str_extract_all(dates, "\\d{2}/\\d{2}/\\d{4}")

```


### <(.+?)>.+?</\\1>
  - This expression conforms to formats similar to html code. Figured this out by removing the quantifiers to understand the structure.
```{r}

code <- c("<div> </div>", "<noscript>...</noscript>", "<div></noscript>")
code

str_extract_all(code, "<(.+?)>.+?</\\1>")

```



## Bonus: Problem 9
The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials at www.r-datacollection.com.

**clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5 fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr**

```{r Bonus}

#Remove all letters except the uppercase ones
bonus <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5 fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

extract <- unlist(str_extract_all(bonus, "[A-Z\\W]+"))
extract #What the letters look like after extracted

result <- str_c(extract, collapse = "") #pastes the individual characters together
result 

result <- str_remove(result, "\\s") #removes the first whitespace
result <- str_remove(result, "\\s") #...second whitespace
result <- gsub("\\.", " ", result) #replaces periods with space where neccessary. 

noquote(result) #Displays final product without quotes

```


