---
title: 'DATA607 WK9 Assignment: APIs'
author: "Javern Wilson"
date: "March 29, 2019"
output: 
  html_document:
    toc: true
    code_folding: hide
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: sandstone
    highlight: kate
---


## Books

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(httr)
library(plyr)
library(kableExtra)
library(jsonlite)
library(tidyverse)
```

```{r}
#get data via API
BOOKS.json <- GET("https://api.nytimes.com/svc/books/v3/lists.json?list=paperback-nonfiction&api-key=3AzC0RPE5QAZ0NAF7rr3cyCPG6sJGwW5")

#parse from JSON into R
non_fiction <- content(BOOKS.json, "parse")

# create a vector with the results from parsing
paper_non_fict <-sapply(non_fiction$results,unlist)
```

When working with JSON and importing to an R object, it usually ends up as a list. Goal is to unlist, transpose, then convert to dataframe. "rbind.fill" will be used to padd the dataframe where data is missing. Based on the imported data there are 15 books.

```{r}
#displays as a dataframe but with the desired column names are the row names
topbooks<- t(do.call("rbind.fill", lapply(lapply(paper_non_fict, t), data.frame, stringsAsFactors = F)))

#so i transposed it a second time
topbooks <- as.data.frame(t(topbooks))

kable(topbooks) %>% kable_styling("striped", "hovered", font_size = 10) %>% scroll_box(height = "500px")



```



## Articles 
This is just to practice more. Just wanted to see if the same code would work for the multi-level variables. There are 235 variables in this one.
```{r}

articles.json <- GET("https://api.nytimes.com/svc/archive/v1/2019/3.json?api-key=3AzC0RPE5QAZ0NAF7rr3cyCPG6sJGwW5")

articles <- content(articles.json, "parse")
y <-sapply(articles$response$docs, unlist)

x <- t(do.call("rbind.fill", lapply(lapply(y, t), data.frame, stringsAsFactors = F)))
x <- as.data.frame(t(x))

kable(head(x, 20)) %>% kable_styling("striped", "hovered", font_size = 11) %>% scroll_box(height = "500px")

```

