---
title: "DATA607 Project 3"
author: "Javern Wilson"
date: "March 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(DBI)
library(RMySQL)
library(RSQLite)
library(tidyverse)
library(kableExtra)
library(gganimate)
library(gifski)
library(plotly)
library(ggmap)
library(leaflet)
```
## Tobacco
As mentioned earlier, Tobacco is one of the top Chronic Disease Indicator of preventable deaths in the US.

```{r}

db_user <- 'newuser'
db_password <- 'pass'
db_name <- 'uscdi'
db_host <- '127.0.0.1' # for access to SQL Database hosted on Google Cloud
db_port <- 3306

drv <- dbDriver("MySQL")

# creating tbdb connection object with RMysql package
connect <- dbConnect(drv, user = db_user, password = db_password, dbname = db_name, host = db_host, port = db_port)

tblTobacco <- dbGetQuery(connect, "SELECT * FROM USChronicDiseaseIndicators WHERE Topic = 'Tobacco' AND LocationAbbr <> 'US';" )
```


```{r}
Tobacco_df <- tblTobacco

dim(Tobacco_df)

kable(head(Tobacco_df, 20)) %>% kable_styling(bootstrap_options = "striped", full_width = T,  font_size = 10) %>%  scroll_box(height = "500px")

```


```{r}
#remove emply columns
Tobacco_df <- Tobacco_df[!sapply(Tobacco_df, function(t) all(t == ""))]
Tobacco_df <- Tobacco_df %>% select("YearEnd", "LocationAbbr", "Question", "DataValue", "Stratification1", "ResponseID") %>% separate(ResponseID, c("Latitude", "Longitude"), sep = ",")

#remove none digit elements from the observations in both longitude and latitude columns and then convert remaining characters to numeric
Tobacco_df$Latitude <- as.numeric(unlist(str_remove(Tobacco_df$Latitude, "^[(]"))) 
Tobacco_df$Longitude <- as.numeric(unlist(str_remove(Tobacco_df$Longitude, "[)]$")))

#rename a few columns
names(Tobacco_df)[c(1, 2, 5)] <- c("Year", "State", "Stratification")

#If any cell for the row is empty then remove the row
Tobacco_df <- Tobacco_df[!Tobacco_df$DataValue == "", ]

```


After making changes to the Tobacco dataframe, here is a preview to a more condensed form
```{r}
dim(Tobacco_df)
kable(head(Tobacco_df, 100)) %>% kable_styling(font_size = 10) %>%  scroll_box(height = "500px")
```

### Questions
These are the type of questions looked at concerning Tobacco
```{r}
Questions <- unique(Tobacco_df$Question)
kable(Questions) %>% kable_styling(font_size = 12) %>%  scroll_box(height = "300px")

```

<br/>A Second Tobacco Dataframe was created because some of the DataValues are categorical (Yes, No) while others are numeric.The first Tobacco Dataframe now has 18681 obs with 7 variables.
```{r}

#separate categorical datavalues from numerical ones and place in a new df
Tobacco_df2 <- subset(Tobacco_df, Tobacco_df$DataValue == c("Yes","No"))
dim(Tobacco_df2)

kable(Tobacco_df2) %>% kable_styling(font_size = 10) %>%  scroll_box(height = "300px")

#remove those placed in Tobacco_df2 from Tobacco_df
new_Tobacco_df <- Tobacco_df[!(Tobacco_df$DataValue %in% Tobacco_df2$DataValue), ]

#convert the datavalue column to numeric

new_Tobacco_df$DataValue <- as.numeric(new_Tobacco_df$DataValue)
new_Tobacco_df$Year <- as.numeric(new_Tobacco_df$Year)

dim(new_Tobacco_df)


```

### Exploration


```{r}

ggplotly(ggplot(new_Tobacco_df, aes(x=State, y=Year, fill=Year)) +
  geom_bar(stat="identity") +  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(axis.text=element_text(size=6)) + ggtitle("Data collected Over the years for States Concerning Tobacco"))

```
<br/> Kansas had a lot to say as most reports came from that state


```{r}

new_Tobacco_df %>% filter(Question == c("Current smokeless tobacco use among youth", "Current cigarette smoking among youth")) %>%  ggplot(aes(x = Year, y = DataValue, fill = Question)) + geom_bar(stat = "identity", position = position_dodge())+ ggtitle("Youth (Grade 9-12) Using Tobacco Over the Years") +  scale_fill_discrete(labels = c("Smokeless Tobacco", "Smoke Tobacco"))


```
<br/>It is likely that if the yout is using smokeless Tobacco, then they'll smoke it too.



```{r}

new_Tobacco_df %>% filter(!Stratification == "Overall") %>%  ggplot(aes(x= Stratification, y=Year, fill=Stratification)) +
  geom_bar(stat="identity") +  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(axis.text=element_text(size=6)) + transition_time(round(Year)) + labs(title = "Year: {frame_time}")

```
Based on this research, most of Tobacco customers are Male and the top race is White.

### Amount of Tobacco Product Excise Tax?

Increasing the price of tabacco discourages people from buying them, especially youths. This way improvements in health will get better in the long term. Majority of the states did not excise tax from the Tobacco product.
```{r}

Tobacco_df2 %>%  filter(Question == "Amount of tobacco product excise tax") %>% ggplot(aes(x = DataValue)) + geom_bar() 

```

### States that allow stronger local tobacco control and prevention laws
<br/> Map below shows which States did and did not allow stronger laws.
```{r}

Tobacco_df2 %>%  filter(Question == "States that allow stronger local tobacco control and prevention laws") %>% ggplot(aes(x = DataValue)) + geom_bar() 

ggmap::register_google(key = "AIzaSyAPKYO8wG3mQUqSL6Ym4HeSIjkZmCwwVP8")
leaflet(options = leafletOptions(minZoom = 2, maxZoom = 8)) %>% 
  addProviderTiles("Esri") %>% 
  setView(lng = -100.166666, lat = 48.166666, zoom = 4) %>% 
  addMarkers(lng =Tobacco_df2$Longitude, lat = Tobacco_df2$Latitude, popup = Tobacco_df2$State, label = Tobacco_df2$DataValue)
```

