---
title: "DATA607 Tidyverse: dplyr"
author: "Javern Wilson"
date: "April 28, 2019"
output: 
  html_document:
    toc: true
    code_folding: show
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: flatly
    highlight: kate
---


#### Create an Example.Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample "vignette" that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.

## Load the library
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```
## Data source
https://www.kaggle.com/ronitf/heart-disease-uci

## Read the data using readr

```{r message=FALSE, warning=FALSE}
heart <- read_csv("https://raw.githubusercontent.com/javernw/DATA607GroupProjects/master/heart.csv")
heart <- as.tibble(heart)
head(heart)
```


The selected package I want to use id dplyr.

## Capability 3.

### `slice` capability tutorial

### Description

Using `slice` we can select rows by specifying the row number.

### Usage
slice(.data, ...)

### Example

To select rows 6 to 12

```{r}
slice(heart, 6:12)
```
<br/>

**OR**

To select rows 10 to 15, 18 and 299 to 302

```{r}
heart %>% slice(c(10:15, 18, 299:302)) #passed as a vector
```

## Capability 4.

### `mutate` capability tutorial

###Description

Creates new columns based on existing ones

### Usage
mutate(.data, ...)

###Example

Let's look at a ratio of resting blood pressure to cholesterol under a new column name "Ratio"
```{r}
#heart %>% mutate(Ratio = trestbps/chol)
heart %>% select(trestbps, chol) %>% mutate(Ratio = trestbps/chol)

```



## Capability 5.

### `summarise` capability tutorial

### Description

Applies functions that return results of length 1. Can perform multiple calculations in the same call.

### Usage
summarise(data, ...)

###Example

Lets see mean and median cholesterol along with mean and median maximum heart rate achieved.

```{r}
heart %>% summarise(Avg_cholesteral = mean(chol), Median_cholesteral = median(chol), Avg_heartrate = mean(thalach), Median_heartrate = median(thalach))
```




## Capability 6.

### `do` capability tutorial (do anything)

### Description

Performs any arbitrary computations on the data

### Usage
do(.data, ...)

###Example

We can create a function that sorts the data by age then returns the first 3 for each age group.

```{r}
x = 3
top <- function(t, x){
  t %>% arrange(desc(age)) %>% head(x)
}
heart %>% group_by(age) %>% do(top(., x))
```
 
 